# My First Knitr Document 
CRPB

## Introduction

-This is some text.  Here is a code chunk.
-command results = "hide" hides the result 
-command echo=FALSE does NOT print the code
-you can name the the code chuck; this is added after you write r.  In the example below, word "simulation" is the name of the code chunk

```{r simulation, echo=FALSE, results="hide"}
set.seed(1)
x <- rnorm(100)
mean(x)
```

- command time <- format(Sys.time() tells you the current time
- if you write a command with a function that needs to be completed such as the one above of time, then this can be processed by stating R and the name of the function and puting it in between of the following  ` ` 

```{r writetime, echo=FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```
```{r}

```

-The Current time is `r time`. My favorite random number is `r rand`.

-Below is an example of an equation and its scatterplot of the data.  This will be split into 2 chunks 

1. Let's first simulate some data  
```{r simulatedata, echo=TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd = 0.5)
```

2. Here is a scatterplot of the data.
```{r scatterplot, fig.height= 4}
par(mar =c(5, 4, 1, 1), las = 1)
plot(x, y, main = "My Simulated Data")
```

-Below is an example making a table using xtable 

A.  Load the data using datasets -> airquality 
```{r fitmodel}
library(datasets)
library(xtable)
data(airquality)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
```
